---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Heartbeat

<!-- badges: start -->
<!-- badges: end -->

The goal of the Heartbeat package is to provide users with a fast and easy method of conducting Heartbeat Analysis on Numeric Likert data with a single function. Heartbeat Analysis itself is a simple method of gauging sentiment of respondents at the item level by using within-group methods to highlight when an individual's response is significantly higher or lower than their own typical baseline.

Additionally, there are several arguments that can be used within this package which are covered more in-depth in the Introduction.rmd file.

## Installation

Installation of the package can be done by copying the line of code below.

``` r
devtools::install_github("PatrickGreen93/Heartbeat")
```

## Example

Below is a basic example of how the code works and a reproducible dataset. 

```{r example, warning=FALSE, message=FALSE}
library(Heartbeat)
# For the purpose of this vignette, I will create a reproducible data.frame
set.seed(43)
n <- 30
y <- 100
example <- data.frame()
for (i in 1:n){
  x <- data.frame(sample(1:5, y, replace = TRUE, prob = c(0.1, 0.2, 0.4, 0.2, 0.1)))
  colnames(x) <- paste0("Q",i)
  example[1:y,i] <- x
}
example[1:5,1:5]

example <- heartbeat(example, type = "counts", threshold = 1, props = FALSE, controversial = TRUE)
example[1:5,]

```
